---
output: github_document
editor_options: 
  chunk_output_type: inline
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# muiDataGrid

<!-- badges: start -->
```{r message=FALSE, warning=FALSE, include=FALSE}
# get JS dependencies (react now lives in devDependencies)
pkg <- jsonlite::fromJSON("js/package.json")
pkgVersion <- pkg$dependencies[["@mui/x-data-grid"]]
reactVersion <- pkg$devDependencies[["react"]]
```

[![CRAN status](https://www.r-pkg.org/badges/version/muiDataGrid)](https://CRAN.R-project.org/package=muiDataGrid)
[![R-CMD-check](https://github.com/lgnbhl/muiDataGrid/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lgnbhl/muiDataGrid/actions/workflows/R-CMD-check.yaml)
`r badger::badge_custom("@mui/x-data-grid", pkgVersion, "blue", "https://mui.com/x/react-data-grid/")`
`r badger::badge_custom("react", reactVersion, "blue", "https://mui.com/x/react-data-grid/")`
[![LinkedIn](https://img.shields.io/badge/LinkedIn-Follow-E4405F?style=social&logo=linkedin)](https://www.linkedin.com/in/FelixLuginbuhl)
<!-- badges: end -->

**muiDataGrid** gives access to [MUI X Data Grid](https://mui.com/x/react-data-grid/), a fast and extensible React data table and React data grid, with filtering, sorting, pagination, and more.

## Install


``` r
install.packages("muiDataGrid")
```

You can install the development version of muiDataGrid like so:

``` r
remotes::install_github("lgnbhl/muiDataGrid")
```

## Basic examples

A minimal example:

```r
library(muiDataGrid)

DataGrid(
  rows = head(iris)
)
```

Customization can be done very easily:

```r
library(muiDataGrid)
library(dplyr)

DataGrid(
  rows = starwars,
  columns = list(
    list(field = "name", headerName = "Names", description = "Names of Starwars character"),
    list(field = "height", headerName = "Height (cm)", description = "Height in centimeter"),
    list(field = "mass", headerName = "Mass (kg)", description = "Mass in kilogram")
  ),
  initialState = list(
    pagination = list(
      paginationModel = list(pageSize = 5)
    )
  ),
  showToolbar = TRUE
)
```

### Resources

- [Package documentation](https://felixluginbuhl.com/muiDataGrid/)
- [All R examples](https://github.com/lgnbhl/muiDataGrid/tree/main/inst/examples
- [Official MUI X Data Grid docs](https://mui.com/x/react-data-grid/)

### Acknowledgements

`muiDataGrid` is built on top of [shiny.react](https://github.com/Appsilon/shiny.react), the R package by [Appsilon](https://www.appsilon.com/) that makes it possible to use React components in Shiny and Quarto.

### Contributing

If you have any issue, question or want to contribute with a pull request, don't hesitate to write me on https://github.com/lgnbhl/muiDataGrid/

Follow [Felix Luginbuhl](https://linkedin.com/in/FelixLuginbuhl) on LinkedIn for updates.

### Trademark notice

"MUI" and "MUI X Data Grid" are trademarks of MUI Inc. `muiDataGrid` is an independent, community-maintained R package and is **not affiliated with, sponsored by, or endorsed by MUI Inc.** The names are used solely to identify the underlying JavaScript library that this package wraps. See MUI's [legal information](https://mui.com/legal/) for the canonical list of MUI's published policies.


