---
output: github_document
editor_options: 
  chunk_output_type: inline
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# muiDataGrid

<!-- badges: start -->
```{r message=FALSE, warning=FALSE, include=FALSE}
# get JS dependencies
dependencies <- jsonlite::fromJSON("js/package.json")$dependencies
```

[![CRAN status](https://www.r-pkg.org/badges/version/muiDataGrid)](https://CRAN.R-project.org/package=muiDataGrid)
[![R-CMD-check](https://github.com/lgnbhl/muiDataGrid/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/lgnbhl/muiDataGrid/actions/workflows/R-CMD-check.yaml)
`r badger::badge_custom(names(dependencies)[4], dependencies[4], "blue", "https://mui.com/x/react-data-grid/")`
`r badger::badge_custom(names(dependencies)[5], dependencies[5], "blue", "https://mui.com/x/react-data-grid/")`
[![LinkedIn](https://img.shields.io/badge/LinkedIn-Follow-E4405F?style=social&logo=linkedin)](https://www.linkedin.com/in/FelixLuginbuhl)
<!-- badges: end -->

**muiDataGrid** gives access to [MUI X Data Grid](https://mui.com/x/react-data-grid/), a fast and extensible React data table and React data grid, with filtering, sorting, aggregation, and more.

## Install

You can install the development version of muiDataGrid like so:

``` r
remotes::install_github("lgnbhl/muiDataGrid")
```

## Basic examples

A minimal example:

```r
library(muiDataGrid)

DataGrid(
  rows = head(iris)
)
```

Customization can be done very easily:

```r
library(muiDataGrid)
library(dplyr)

DataGrid(
  rows = starwars,
  columns = data.frame(
    field = c("name", "height", "mass"), # column names
    headerName = c("Names", "Height (cm)", "Mass (kg)"),
    description = c("Names of Starwars character", "Height in centimeter", "Mass in kilogram")
  ),
  initialState = list(
    pagination = list(
      paginationModel = list(pageSize = 5)
    )
  ),
  showToolbar = TRUE
)
```

**Read the full documentation with examples [here](https://felixluginbuhl.com/muiDataGrid/).**

### Contribute

If you have any issue, question or want to contribute with a pull request, don't hesitate to write me on https://felixluginbuhl.com/muiDataGrid/

For updates follow [Felix Luginbuhl](https://linkedin.com/in/FelixLuginbuhl) on LinkedIn.
