---
title: "Column Dimensions"
format:
  html:
    minimal: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Column Dimensions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(muiDataGrid)
library(dplyr)
```

Column dimensions control how columns are sized within the data grid. The grid provides flexible options for both fixed and dynamic column widths.

## Width

### Fixed Width

The simplest way to control column width is to set a fixed pixel value:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", width = 250),
    list(field = "height", width = 100),
    list(field = "mass", width = 100)
  )
)
```

### Default Width

If no width is specified, columns default to 100 pixels wide.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head()
  # All columns will be 100px wide
)
```

## Fluid Width

Use the `flex` property to make columns grow proportionally to fill available space. The `flex` value determines the proportion of available space each column takes.

### Basic Flex

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", flex = 1),
    list(field = "height", flex = 1),
    list(field = "mass", flex = 1)
  )
)
```

All columns share the available space equally.

### Proportional Flex

Columns with higher flex values take proportionally more space:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, hair_color) |> head(),
  columns = list(
    list(field = "name", flex = 3),        # Takes 3/5 of flexible space
    list(field = "height", flex = 1),      # Takes 1/5 of flexible space
    list(field = "mass", flex = 1),        # Takes 1/5 of flexible space
    list(field = "hair_color", width = 150) # Fixed width, not flexible
  )
)
```

### Combining Fixed and Flex

You can mix fixed width columns with flex columns:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, hair_color) |> head(),
  columns = list(
    list(field = "name", flex = 2),
    list(field = "height", width = 100),    # Fixed
    list(field = "mass", width = 100),      # Fixed
    list(field = "hair_color", flex = 1)
  )
)
```

The fixed columns maintain their width, while flex columns share the remaining space.

## Minimum Width

Set a minimum width constraint to prevent columns from becoming too narrow:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", flex = 1, minWidth = 150),
    list(field = "height", flex = 1, minWidth = 80),
    list(field = "mass", flex = 1, minWidth = 80)
  )
)
```

When the grid is resized smaller, columns will not shrink below their `minWidth`.

## Maximum Width

Set a maximum width constraint to prevent columns from becoming too wide:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", flex = 1, maxWidth = 200),
    list(field = "height", flex = 1),
    list(field = "mass", flex = 1)
  )
)
```

## Min and Max Width Together

Combine minimum and maximum width for precise control:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", flex = 1, minWidth = 120, maxWidth = 300),
    list(field = "height", width = 100),
    list(field = "mass", width = 100)
  )
)
```

This ensures columns stay within a reasonable range while still being flexible.

## Column Resizing

### Resizable Columns

By default, users can manually resize columns by dragging the column header edges:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head()
  # Column resizing enabled by default
)
```

### Disable Resizing Per Column

Control which columns can be resized:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", resizable = TRUE),
    list(field = "height", resizable = FALSE), # Cannot resize
    list(field = "mass", resizable = FALSE)    # Cannot resize
  )
)
```

### Disable Resizing Globally

Disable column resizing for the entire grid:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  disableColumnResize = TRUE
)
```

## Autosizing Columns

You can programmatically auto-size columns to fit their content. This is typically done through the grid's API, but you can set initial widths based on content length:

```{r}
# Estimate widths based on content
DataGrid(
  rows = starwars |> select(name, height, mass, hair_color) |> head(),
  columns = list(
    list(field = "name", width = 180),        # Longer names
    list(field = "height", width = 90),       # Short numbers
    list(field = "hair_color", width = 120)   # Medium text
  )
)
```
