---
title: "Column Definition"
format:
  html:
    minimal: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Column Definition}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(muiDataGrid)
library(dplyr)
```

# Column Definition

Column definitions in the Data Grid specify how data is displayed, formatted, and interacted with. Each column has a unique field name and various properties that control its behavior.

## Core Properties

### field

The unique identifier for the column (required). Must match a field in the row data.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name"),
    list(field = "height"),
    list(field = "mass")
  )
)
```

### headerName

The text displayed in the column header. Defaults to the field name if not specified.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", headerName = "Character Name"),
    list(field = "height", headerName = "Height (cm)"),
    list(field = "mass", headerName = "Weight (kg)")
  )
)
```

### description

Tooltip text shown when hovering over the column header.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(
      field = "name",
      headerName = "Name",
      description = "The character's full name"
    ),
    list(
      field = "height",
      headerName = "Height",
      description = "Height in centimeters"
    ),
    list(
      field = "mass",
      headerName = "Mass",
      description = "Weight in kilograms"
    )
  )
)
```

### type

The column type affects formatting and default behavior. Available types:
- `"string"` (default)
- `"number"`
- `"date"`
- `"dateTime"`
- `"boolean"`
- `"singleSelect"`

```{r}
starwars_typed <- starwars |>
  select(name, height, birth_year) |>
  mutate(is_tall = height > 180) |>
  head()

DataGrid(
  rows = starwars_typed,
  columns = list(
    list(field = "name", type = "string", headerName = "Name", width = 200),
    list(field = "height", type = "number", headerName = "Height", width = 120),
    list(field = "birth_year", type = "number", headerName = "Birth Year", width = 120),
    list(field = "is_tall", type = "boolean", headerName = "Tall?", width = 100)
  )
)
```

## Display Properties

### width

Fixed column width in pixels.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", width = 250),
    list(field = "height", width = 120),
    list(field = "mass", width = 120)
  )
)
```

### flex

Flexible column width as a proportion of available space. Cannot be used with `width`.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", flex = 2),
    list(field = "height", flex = 1),
    list(field = "mass", flex = 1)
  )
)
```

### minWidth / maxWidth

Constraints for column width when using `flex` or resizing.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", flex = 1, minWidth = 150),
    list(field = "height", flex = 1, minWidth = 80, maxWidth = 120),
    list(field = "mass", flex = 1, minWidth = 80, maxWidth = 120)
  )
)
```

### align / headerAlign

Text alignment for cell content and header.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(
      field = "name", 
      headerAlign = "left", 
      align = "left", 
      width = 200
    ),
    list(
      field = "height", 
      headerAlign = "center", 
      align = "right", 
      width = 120
    ),
    list(
      field = "mass", 
      headerAlign = "center", 
      align = "right", 
      width = 120
    )
  )
)
```

### hide

Hide a column while keeping it in the data.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", width = 200),
    list(field = "height", width = 120),
    list(field = "mass", width = 120, hide = TRUE),
    list(field = "species", width = 150)
  )
)
```

## Interaction Properties

### sortable

Enable or disable sorting for the column (default: `TRUE`).

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", sortable = TRUE, width = 200),
    list(field = "height", sortable = TRUE, width = 120),
    list(field = "mass", sortable = FALSE, width = 120)
  )
)
```

### filterable

Enable or disable filtering for the column (default: `TRUE`).

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", filterable = TRUE, width = 200),
    list(field = "height", filterable = TRUE, width = 120),
    list(field = "mass", filterable = FALSE, width = 150)
  )
)
```

### editable

Make the column editable (default: `FALSE`).

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", editable = FALSE, width = 200),
    list(field = "height", editable = TRUE, width = 120),
    list(field = "mass", editable = TRUE, width = 120)
  )
)
```

### resizable

Enable column resizing by dragging borders (default: `TRUE`).

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", resizable = TRUE, width = 200),
    list(field = "height", resizable = TRUE, width = 120),
    list(field = "mass", resizable = FALSE, width = 120)
  )
)
```

## Value Formatting

### valueFormatter

Format the displayed value without changing the underlying data.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", width = 200),
    list(
      field = "height",
      headerName = "Height",
      width = 150,
      valueFormatter = JS("function(value) { return value ? value + ' cm' : 'Unknown'; }")
    ),
    list(
      field = "mass",
      headerName = "Mass",
      width = 150,
      valueFormatter = JS("function(value) { return value ? value + ' kg' : 'Unknown'; }")
    )
  )
)
```

### valueGetter

Compute or derive values from row data.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name", width = 150),
    list(field = "height", width = 100),
    list(field = "mass", width = 100),
    list(
      field = "bmi",
      headerName = "BMI",
      width = 100,
      type = "number",
      valueGetter = JS("
        function(value, row) {
          if (row.height && row.mass) {
            const heightInM = row.height / 100;
            return Math.round(row.mass / (heightInM * heightInM) * 10) / 10;
          }
          return null;
        }
      ")
    )
  )
)
```

### valueSetter

Control how edited values are stored (for editable columns):

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(field = "name"),
    list(
      field = "height",
      editable = TRUE,
      valueSetter = JS("
        function(value, row) {
          const numValue = Number(value);
          if (!isNaN(numValue) && numValue > 0) {
            row.height = numValue;
            return row;
          }
          return row; // Return unchanged if invalid
        }
      ")
    ),
    list(field = "mass")
  )
)
```

## Advanced Properties

### renderCell

Custom cell rendering with React components:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(
      field = "name",
      renderCell = JS("
        function(params) {
          return React.createElement('strong', {}, params.value);
        }
      ")
    ),
    list(field = "height"),
    list(field = "mass")
  )
)
```

### renderHeader

Custom header rendering:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(
      field = "name",
      renderHeader = JS("
        function(params) {
          return React.createElement('strong', {}, params.colDef.headerName || params.field);
        }
      ")
    ),
    list(field = "height"),
    list(field = "mass")
  )
)
```

## Column Grouping

Group multiple columns under a common header:

```{r}
DataGrid(
  rows = starwars |> 
    select(name, height, mass, hair_color, skin_color) |> 
    head(),
  columns = list(
    list(field = "name"),
    list(field = "height"),
    list(field = "mass"),
    list(field = "hair_color"),
    list(field = "skin_color")
  ),
  columnGroupingModel = list(
    list(
      groupId = "physical",
      headerName = "Physical Attributes",
      children = list(
        list(field = "height"),
        list(field = "mass")
      )
    ),
    list(
      groupId = "appearance",
      headerName = "Appearance",
      children = list(
        list(field = "hair_color"),
        list(field = "skin_color")
      )
    )
  )
)
```

## Complete Example

Here's a comprehensive example using many column properties:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, hair_color, skin_color) |> head(10),
  columns = list(
    list(
      field = "name",
      headerName = "Character Name",
      description = "Name of the Star Wars character",
      width = 180,
      sortable = TRUE,
      filterable = TRUE,
      hideable = FALSE,
      pinnable = TRUE,
      align = "left"
    ),
    list(
      field = "height",
      headerName = "Height",
      description = "Height in centimeters",
      type = "number",
      width = 100,
      align = "center",
      headerAlign = "center",
      valueFormatter = JS("function(value) { return value ? value + ' cm' : 'N/A'; }")
    ),
    list(
      field = "mass",
      headerName = "Mass",
      description = "Mass in kilograms",
      type = "number",
      width = 100,
      align = "center",
      headerAlign = "center",
      valueFormatter = JS("function(value) { return value ? value + ' kg' : 'N/A'; }")
    )
  ),
  initialState = list(
    pagination = list(
      paginationModel = list(pageSize = 5)
    )
  )
)
```
