---
title: "Custom with React JS"
format:
  html:
    minimal: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Custom with React JS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(muiDataGrid)
library(dplyr)
```

## Custom columns using Javascript

Using `valueGetter` or `valueGetter`:

```{r}
# https://stackoverflow.com/a/79508826
library(dplyr)

# "id" column is mandatory
df <- starwars |>
  select(1:2) |>
  mutate(
    id = row_number()
)

DataGrid(
  sx = list(height = 300),
  rows = df,
  columns = list(
    list(
      field = "name"
    ),
    list(
      field = "height",
      type = "number",
      headerName = "Height",
      description = "Height in meters",
      valueGetter = JS("
        function(value) {
          if (value == null) { return ''; }
          return value / 100 }"
      ),
      valueFormatter = JS("
        function(value) {
          if (value == '') { return ''; }
          return `${value} cm`}"
      )
    )
  )
)
```

Using `valueGetter` and `renderCell`:

```{r}
library(shiny.react)

# "id" column is mandatory
df <- starwars |>
  select(1:2) |>
  mutate(
    id = row_number()
)

DataGrid(
  sx = list(height = 300),
  rows = df,
  columns = list(
    list(
      field = "name"
    ),
    list(
      field = "height",
      valueGetter = JS("function(value) { return value / 100 }"),
      renderCell = JS("function(params) { return `${params.value} m` }")
    )
  )
)
```

## Custom table using React JS Components

If you don't know React, I would suggest to ask an LLM to create a React function for you. All the examples below have been created using the help of an LLM:

Below an simple example to add a custom image in your table (i.e. wrapping the value in an styled "img" HTML tag):

```{r}
library(shiny.react)

df <- data.frame(
  country = c("Switzerland", "France"),
  img = c("https://flagsapi.com/CH/flat/64.png", "https://flagsapi.com/FR/flat/64.png")
)

DataGrid(
  sx = list(height = 300),
  rows = df,
  columns = list(
    list(
      field = "img",
      headerName = "",
      width = 10,
      sortable = FALSE,
      disableColumnMenu = TRUE,
      renderCell = JS("
        function(params) { 
          return React.createElement(
            'img',
            {
              src: params.value,
              alt: 'flag',
              style: {
                width: '100%',
                height: '100%',
                objectFit: 'contain',
                display: 'block'
              }
            }
          );
        }"
      )
    ),
    list(
      field = "country"
    )
  )
)
```

You can create an React function to color the values and cell background following a logic:

```{r}
df <- data.frame(
  id = c(1,2,3,4,5),
  value = c(1,2,3,4,5)
)

DataGrid(
  rows = df,
  columns = list(
    list(
      field = "value",
      renderCell = JS("
        function(params) { 
        // Colored Value
        const ColoredValue = ({ value }) => {
          let bgColor = '#e8f5e9'; // light green
          let textColor = '#2e7d32';
          
          if (value < 2) {
            bgColor = '#ffebee'; // light red
            textColor = '#c62828';
          } else if (value < 3) {
            bgColor = '#ffe0b2'; // light orange
            textColor = '#e65100';
          } else if (value < 4) {
            bgColor = '#fff9c4'; // light yellow
            textColor = '#f57f17';
          }
          return React.createElement(
            'div',
            {
              style: {
                width: '100%',
                height: '100%',
                display: 'flex',
                alignItems: 'center',
                justifyContent: 'center',
                backgroundColor: bgColor,
                color: textColor,
                fontWeight: '600',
                fontSize: '13px',
                margin: '-8px -16px', // Negative margin to counteract cell padding
                padding: '8px 16px'   // Add padding back to maintain text position
              }
            },
            value.toFixed(1)
          );
        };
        return React.createElement(ColoredValue, { value: params.value })
        }")
    )
  )
)
```

You can also use [Google Material Icons](https://fonts.google.com/icons) by loading the font stylesheet with `htmltools::tags$link()`:

```{r}
df <- data.frame(
  id = c(1,2,3,4,5),
  value = c(1,2,3,4,5)
)

htmltools::tagList(
  htmltools::tags$link(
    rel = "stylesheet",
    href = "https://fonts.googleapis.com/icon?family=Material+Icons+Outlined"
  ),
  DataGrid(
    rows = df,
    columns = list(
      list(
        field = "value",
        renderCell = JS("
          function(params) {
          const ColoredValue = ({ value }) => {
            let bgColor = '#e8f5e9'; // light green
            let textColor = '#2e7d32';
            let icon = 'sentiment_very_satisfied'; // very happy

            if (value < 2) {
              bgColor = '#ffebee'; // light red
              textColor = '#c62828';
              icon = 'sentiment_very_dissatisfied'; // very sad
            } else if (value < 3) {
              bgColor = '#ffe0b2'; // light orange
              textColor = '#e65100';
              icon = 'sentiment_dissatisfied'; // dissatisfied
            } else if (value < 4) {
              bgColor = '#fff9c4'; // light yellow
              textColor = '#f57f17';
              icon = 'sentiment_satisfied'; // satisfied
            }

            return React.createElement(
              'div',
              {
                style: {
                  width: '100%',
                  height: '100%',
                  display: 'flex',
                  alignItems: 'center',
                  justifyContent: 'center',
                  gap: '8px',
                  backgroundColor: bgColor,
                  color: textColor,
                  fontWeight: '600',
                  fontSize: '13px',
                  margin: '-8px -16px',
                  padding: '8px 16px'
                }
              },
              React.createElement('span', {
                className: 'material-icons-outlined',
                style: { fontSize: '18px' }
              }, icon),
              value.toFixed(1)
            );
          };
          return React.createElement(ColoredValue, { value: params.value })
          }")
      )
    )
  )
) |>
  htmltools::browsable() # show in viewer
```

You can add a React function to create a custom progress bar:

```{r}
df <- data.frame(
  id = c(1, 2, 3),
  value = c(20, 60, 80)
)

DataGrid(
  rows = df,
  columns = list(
    list(
      field = "value",
      renderCell = JS("
        function(params) { 
        // Progress Bar Component
        const ProgressBar = ({ value }) => {
          return React.createElement(
            'div',
            {
              style: {
                width: '100%',
                height: '100%',
                display: 'flex',
                alignItems: 'center'
              }
            },
            React.createElement(
              'div',
              {
                style: {
                  width: '100%',
                  height: '20px',
                  backgroundColor: '#e0e0e0',
                  overflow: 'hidden',
                  position: 'relative'
                }
              },
              React.createElement('div', {
                style: {
                  width: `${value}%`,
                  height: '100%',
                  backgroundColor: value < 30 ? '#f44336' : value < 70 ? '#ff9800' : '#4caf50',
                  transition: 'width 0.3s ease'
                }
              }),
              React.createElement(
                'span',
                {
                  style: {
                    position: 'absolute',
                    top: '50%',
                    left: '50%',
                    transform: 'translate(-50%, -50%)',
                    fontSize: '11px',
                    fontWeight: '600',
                    color: '#333',
                    textShadow: '0 0 2px rgba(255,255,255,0.8)'
                  }
                },
                `${value}%`
              )
            )
          );
        };
        return React.createElement(ProgressBar, { value: params.value })
        }")
    )
  )
)
```

Here another example using React with a Rating component:

```{r}
df <- data.frame(
  id = c(1, 2, 3),
  value = c(1, 3, 6)
)

DataGrid(
  rows = df,
  columns = list(
    list(
      field = "value",
      renderCell = JS("
        function(params) { 
        // Rating Component
        const Rating = ({ value }) => {
          const maxStars = 5;
          const filledStars = Math.max(1, Math.min(5, Math.round(value))); // Clamp value between 1-5
          
          return React.createElement(
            'div',
            {
              style: {
                width: '100%',
                height: '100%',
                display: 'flex',
                alignItems: 'center',
                justifyContent: 'center',
                gap: '4px'
              }
            },
            ...Array.from({ length: maxStars }, (_, index) => {
              const isFilled = index < filledStars;
              return React.createElement(
                'span',
                {
                  key: index,
                  style: {
                    fontSize: '18px',
                    color: isFilled ? '#ffc107' : '#e0e0e0'
                  }
                },
                '★'
              );
            })
          );
        };
        return React.createElement(Rating, { value: params.value })
        }")
    )
  )
)
```

Once again, just ask an LLM to create a React function according to your needs.
