---
title: "Layout"
format:
  html:
    minimal: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Layout}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
library(muiDataGrid)
library(dplyr)
```

# Layout

The Data Grid offers multiple options to control its layout.

## Height

### Fixed Height

You can specify a fixed height using CSS or the `sx` prop:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(10),
  sx = list(height = 300)
)
```

### Auto Height

The Data Grid can automatically adjust its height to fit the content by using the `autoHeight` prop:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(10),
  autoHeight = TRUE
)
```


## Width

### Fluid Width

By default, the Data Grid takes the full width of its container. The columns flex to fill the available space.

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, species) |> head(10)
)
```

### Fixed Width

You can specify a fixed width using CSS or the `sx` prop:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(10),
  sx = list(width = 600)
)
```

## Pagination

You can add pagination and adjust the page size using `initialState` (here 5 rows per page):

```{r}
library(dplyr)

DataGrid(
  rows = starwars |> select(name, height, mass) |> head(10),
  initialState = list(
    pagination = list(
      paginationModel = list(pageSize = 5)
    )
  )
)
```
