---
title: "Row Definition"
format:
  html:
    minimal: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Row Definition}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message=FALSE, warning=FALSE}
library(muiDataGrid)
library(dplyr)
```

# Row Definition

Rows are the fundamental data units in the Data Grid. This guide covers how to feed data to the grid and work with row identifiers.

## Feeding Data to the Grid

### Basic Data Loading

The grid automatically handles R data frames and tibbles through the `rows` prop:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, species) |> head()
)
```

## Row Identifier

### Automatic ID Generation

The **muiDataGrid** package automatically generates unique IDs for your rows, so you don't need to create an `id` column manually:

```{r}
# No need to add id column - it's handled automatically
DataGrid(
  rows = starwars |> select(name, height, mass) |> head()
)
```

### Pagination

For large datasets, consider using pagination:

```{r}
# Show a subset of data
DataGrid(
  rows = starwars |> head(10),
  initialState = list(
    pagination = list(
      paginationModel = list(pageSize = 5)
    )
  )
)
```
