---
title: "Styling Basics"
format:
  html:
    minimal: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Styling Basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(muiDataGrid)
library(dplyr)
```

# Styling Basics

The Data Grid CSS can be easily customized using various approaches. This guide covers the fundamental styling methods available.

## Using the sx Prop

For one-off styles, the `sx` prop can be used. It lets you apply both simple and complex customizations directly to the Data Grid element. The keys accepted can be any CSS property as well as custom properties provided by MUI.

### Basic Styling

Apply simple styles like margin and padding:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, species) |> head(),
  sx = list(
    m = 2  # margin: 2 * 8px = 16px
  )
)
```

### Advanced Styling with Borders and Shadows

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, species) |> head(),
  sx = list(
    boxShadow = 2,
    border = 2,
    borderColor = "primary.light",
    "& .MuiDataGrid-cell:hover" = list(
      color = "primary.main"
    )
  )
)
```

### Custom Background and Typography

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  sx = list(
    backgroundColor = "#f5f5f5",
    "& .MuiDataGrid-columnHeaders" = list(
      color = "blue",
      fontSize = 20
    ),
    "& .MuiDataGrid-cell" = list(
      borderBottom = "1px solid #e0e0e0"
    )
  )
)
```

## Styling Column Headers

The column definition provides properties to apply styles to column headers:

### headerClassName

Apply custom class names to column headers:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(
      field = "name",
      headerName = "Character Name",
      headerClassName = "custom-header",
      width = 200
    ),
    list(
      field = "height",
      headerName = "Height (cm)",
      headerClassName = "custom-header",
      width = 120
    ),
    list(
      field = "mass",
      headerName = "Mass (kg)",
      headerClassName = "custom-header",
      width = 120
    )
  ),
  sx = list(
    "& .custom-header" = list(
      backgroundColor = "#f0f0f0",
      fontWeight = "bold"
    )
  )
)
```

### headerAlign

Align the content of column headers:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  columns = list(
    list(
      field = "name",
      headerName = "Character",
      headerAlign = "left",
      width = 200
    ),
    list(
      field = "height",
      headerName = "Height",
      headerAlign = "center",
      width = 120
    ),
    list(
      field = "mass",
      headerName = "Mass",
      headerAlign = "right",
      width = 120
    )
  )
)
```

## Styling Rows

### Using getRowClassName

The `getRowClassName` prop applies custom CSS classes to rows based on their data:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, species) |> head(),
  getRowClassName = JS("
    function(params) {
      if (params.row.species === 'Droid') {
        return 'droid-row';
      } else if (params.row.species === 'Human') {
        return 'human-row';
      }
      return '';
    }
  "),
  sx = list(
    "& .droid-row" = list(
      backgroundColor = "#e3f2fd",
      "&:hover" = list(
        backgroundColor = "#bbdefb"
      )
    ),
    "& .human-row" = list(
      backgroundColor = "#fff3e0",
      "&:hover" = list(
        backgroundColor = "#ffe0b2"
      )
    )
  )
)
```

### Conditional Row Styling

Style rows based on numeric values:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> head(),
  getRowClassName = JS("
    function(params) {
      if (params.row.height > 190) {
        return 'tall-character';
      } else if (params.row.height < 100) {
        return 'short-character';
      }
      return '';
    }
  "),
  sx = list(
    "& .tall-character" = list(
      backgroundColor = "#c8e6c9",
      fontWeight = "bold"
    ),
    "& .short-character" = list(
      backgroundColor = "#ffccbc"
    )
  )
)
```

## Styling Cells

### Using cellClassName in Column Definition

Apply CSS classes to all cells in a specific column:

```{r}
DataGrid(
  rows = starwars |> 
    select(name, height, mass) |>
    mutate(bmi = mass / (height/100)^2) |>
    head(15),
  columns = list(
    list(
      field = "name",
      headerName = "Name",
      width = 180,
      cellClassName = "name-cell"
    ),
    list(
      field = "height",
      headerName = "Height",
      width = 100
    ),
    list(
      field = "mass",
      headerName = "Mass",
      width = 100
    ),
    list(
      field = "bmi",
      headerName = "BMI",
      width = 100,
      cellClassName = JS("
        function(params) {
          if (params.value == null) return '';
          if (params.value > 30) return 'high-bmi';
          if (params.value < 18.5) return 'low-bmi';
          return 'normal-bmi';
        }
      ")
    )
  ),
  sx = list(
    "& .name-cell" = list(
      fontWeight = "500"
    ),
    "& .high-bmi" = list(
      backgroundColor = "#ffcdd2",
      color = "#c62828"
    ),
    "& .low-bmi" = list(
      backgroundColor = "#fff9c4",
      color = "#f57f17"
    ),
    "& .normal-bmi" = list(
      backgroundColor = "#c8e6c9",
      color = "#2e7d32"
    )
  )
)
```

### Using getCellClassName at Grid Level

Apply conditional styling to cells across all columns:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass) |> filter(is.na(mass)) |> head(),
  getCellClassName = JS("
    function(params) {
      if (params.field === 'name') return '';
      if (params.value == null) return 'missing-value';
      return '';
    }
  "),
  sx = list(
    "& .missing-value" = list(
      backgroundColor = "lightblue",
      fontStyle = "italic"
    )
  )
)
```

## Cell Alignment

Use the `align` property to change the alignment of cell content:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, species) |> head(),
  columns = list(
    list(
      field = "name",
      headerName = "Character",
      align = "left",
      headerAlign = "left",
      width = 180
    ),
    list(
      field = "height",
      headerName = "Height",
      align = "center",
      headerAlign = "center",
      width = 120
    ),
    list(
      field = "mass",
      headerName = "Mass",
      align = "right",
      headerAlign = "right",
      width = 120
    ),
    list(
      field = "species",
      headerName = "Species",
      align = "center",
      headerAlign = "center",
      width = 150
    )
  )
)
```

## Striped Rows

Create alternating row colors using `indexRelativeToCurrentPage`:

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, species) |> head(),
  getRowClassName = JS("
    function(params) {
      return params.indexRelativeToCurrentPage % 2 === 0 ? 'even' : 'odd';
    }
  "),
  sx = list(
    "& .even" = list(
      backgroundColor = "#f5f5f5"
    ),
    "& .odd" = list(
      backgroundColor = "white"
    )
  )
)
```

### Striped Rows with Hover Effect

```{r}
DataGrid(
  rows = starwars |> select(name, height, mass, homeworld) |> head(),
  getRowClassName = JS("
    function(params) {
      return params.indexRelativeToCurrentPage % 2 === 0 ? 'even' : 'odd';
    }
  "),
  sx = list(
    "& .even" = list(
      backgroundColor = "#fafafa",
      "&:hover" = list(
        backgroundColor = "#eeeeee"
      )
    ),
    "& .odd" = list(
      backgroundColor = "white",
      "&:hover" = list(
        backgroundColor = "#f5f5f5"
      )
    )
  )
)
```

## Complete Styling Example

Here's a comprehensive example combining multiple styling techniques:

```{r}
starwars_styled <- starwars |>
  select(name, height, mass, species, homeworld) |>
  head()

DataGrid(
  rows = starwars_styled,
  columns = list(
    list(
      field = "name",
      headerName = "Character",
      width = 180,
      cellClassName = "bold-cell",
      headerAlign = "left"
    ),
    list(
      field = "height",
      headerName = "Height (cm)",
      width = 120,
      type = "number",
      align = "right",
      headerAlign = "right"
    ),
    list(
      field = "mass",
      headerName = "Mass (kg)",
      width = 120,
      type = "number",
      align = "right",
      headerAlign = "right"
    ),
    list(
      field = "species",
      headerName = "Species",
      width = 130,
      align = "center",
      headerAlign = "center"
    ),
    list(
      field = "homeworld",
      headerName = "Homeworld",
      width = 140
    )
  ),
  getRowClassName = JS("
    function(params) {
      const baseClass = params.indexRelativeToCurrentPage % 2 === 0 ? 'even' : 'odd';
      if (params.row.species === 'Droid') {
        return baseClass + ' droid';
      }
      return baseClass;
    }
  "),
  sx = list(
    border = 2,
    borderColor = "primary.light",
    borderRadius = 2,
    "& .MuiDataGrid-columnHeaders" = list(
      backgroundColor = "#1976d2",
      color = "darkgreen",
      fontSize = 14,
      fontWeight = "bold"
    ),
    "& .bold-cell" = list(
      fontWeight = "600"
    ),
    "& .even" = list(
      backgroundColor = "#fafafa"
    ),
    "& .odd" = list(
      backgroundColor = "white"
    ),
    "& .droid" = list(
      backgroundColor = "#e3f2fd !important"
    ),
    "& .MuiDataGrid-cell:hover" = list(
      color = "primary.main"
    ),
    "& .MuiDataGrid-row:hover" = list(
      backgroundColor = "#f5f5f5"
    )
  )
)
```
